Single-step Hybrid Predition

### 2016-12-07 reports/intro.Rmd
Update equation numbering after updates to bookdown package

$$
\begin{equation}
	\mathbf{y} = \mathbf{Xb} + \mathbf{Za} + \mathbf{e},
	\label{eq:animal_model}
\end{equation}
$$

@@

$$
\begin{equation}
	\label{eq:gebv}
	\mathbf{\tilde{g}} =  (\mathbf{M} - 2\mathbf{p})\mathbf{\hat{u}},
\end{equation}
$$

@@

## Single Step Transciptomic prediction of Breeding Values
The breeding values in a mRNA prediction model are in general given by:
$$
\begin{equation}
	\label{eq:mrnaebv}
	\mathbf{\tilde{g}} = \mathbf{W}\boldsymbol{\hat{\alpha}},
\end{equation}
$$

@@

Consider now the situation in which only a subset of the population has 
transcriptomic information but all have SNP genotypes.
We can take a similar route as in @fernando_class_2014 and impute missing
mRNA data.
The difference here is that &mdash; instead of predicting SNP covariates using 
pedigree relationship &mdash; we impute mRNA covariates using SNP or genomic 
relationships.
Let the subscript $1$ denote individuals for which SNP genotypes but no mRNA 
data are available.
Individuals with subscript $2$ have both, SNP and mRNA information.
The covariates in $\mathbf{W}$ are centered.
The vector $\mathbf{g}_1$ can be written as the sum of the conditional 
expectation given $\mathbf{g}_2$ and a residual error term:
$$
\begin{align}
	\label{eq:mrna1}
	\mathbf{g_1} &= E(\mathbf{g}_1|\mathbf{g}_2) + \boldsymbol{\epsilon} \\
	&= \mathbf{G_{12}}\mathbf{G_{22}}^{-1}\mathbf{W_2}\boldsymbol{\hat{\alpha}} + (\mathbf{g_1} - \mathbf{G_{12}}\mathbf{G_{22}}^{-1}\mathbf{W_2}\boldsymbol{\hat{\alpha}}) \\
	&= \mathbf{\hat{g}}_1 + \boldsymbol{\epsilon}
\end{align}
$$

@@

$$
\begin{equation}
\mathbf{y} = \mathbf{Xb} + \mathbf{W} \boldsymbol{\alpha} + \mathbf{U}_{Ped} \boldsymbol{\epsilon}_{Ped} 
+ \mathbf{U}_{SNP} \boldsymbol{\epsilon}_{SNP}+ \mathbf{e},
\end{equation}
$$

@@

$$
\begin{equation}
\mathbf{X} = 
\begin{bmatrix}
  X_0 \\
  X_1 \\
  X_2 
 \end{bmatrix},
 \mathbf{W} = 
\begin{bmatrix}
  Z_0\hat{\mathbf{W}_0} \\
  Z_1\hat{\mathbf{W}_1} \\
  \mathbf{W}_2 
 \end{bmatrix},
 \mathbf{U}_{Ped} = 
\begin{bmatrix}
  Z_0 \\
  0 \\
  0 
 \end{bmatrix},
 \mathbf{U}_{SNP} = 
\begin{bmatrix}
  0 \\
  Z_1 \\
  0 
 \end{bmatrix},
 \end{equation}
$$

@@

$$
\begin{align}
\begin{bmatrix}
  y_0 \\
  y_1 \\
  y_2 
 \end{bmatrix}
& =
 \begin{bmatrix}
  X_0 \\
  X_1 \\
  X_2 
 \end{bmatrix}
 \boldsymbol{\beta} + 
 \begin{bmatrix}
  Z_0 & 0 & 0 \\
  0 & Z_1 & 0 \\
  0 & 0 & Z_2 
 \end{bmatrix}
\begin{bmatrix}
  g_0 \\
  g_1 \\
  g_2 
 \end{bmatrix}
  + \mathbf{e} \\
    & = 
 \begin{bmatrix}
  X_0 \\
  X_1 \\
  X_2 
 \end{bmatrix}
 \boldsymbol{\beta} + 
 \begin{bmatrix}
  Z_0 & 0 & 0 \\
  0 & Z_1 & 0 \\
  0 & 0 & Z_2 
 \end{bmatrix}
\begin{bmatrix}
  \mathbf{A}_{02}\mathbf{A}_{22}^{-1}\mathbf{W}_2\boldsymbol{\alpha} + \boldsymbol{\epsilon}_{Ped} \\
  \mathbf{G}_{12}\mathbf{G}_{22}^{-1}\mathbf{W}_2\boldsymbol{\alpha} + \boldsymbol{\epsilon}_{SNP}  \\
  \mathbf{W}_2\boldsymbol{\alpha} \\
 \end{bmatrix}
  + \mathbf{e} \\
    & = 
 \begin{bmatrix}
  X_0 \\
  X_1 \\
  X_2 
 \end{bmatrix}
 \boldsymbol{\beta} + 
 \begin{bmatrix}
  Z_0 & 0 & 0 \\
  0 & Z_1 & 0 \\
  0 & 0 & Z_2 
 \end{bmatrix}
\begin{bmatrix}
  \hat{\mathbf{W}_0}\boldsymbol{\alpha} + \boldsymbol{\epsilon}_{Ped} \\
  \hat{\mathbf{W}_1}\boldsymbol{\alpha} + \boldsymbol{\epsilon}_{SNP} \\
  \mathbf{W}_2\boldsymbol{\alpha} \\
 \end{bmatrix}
  + \mathbf{e} \\
\end{align}
$$

@@

$$
\begin{equation}
\tilde{\mathbf{g}} = 
 \begin{bmatrix}
  \hat{\mathbf{W}_0} \\
  \hat{\mathbf{W}_1} \\
  \mathbf{W}_2 
 \end{bmatrix}
 \hat{\boldsymbol{\alpha}}
 + 
 \begin{bmatrix}
  \mathbf{Z}_0 \\
  0 \\
  0
 \end{bmatrix}
 \hat{\boldsymbol{\epsilon}}_{Ped}
 + 
 \begin{bmatrix}
  0 \\
  \mathbf{Z}_1 \\
  0
 \end{bmatrix}
 \hat{\boldsymbol{\epsilon}}_{SNP} 
 \end{equation}
 $$

@@

$$
\begin{align}
\mathbf{y} &= \mathbf{Xb} \\ 
  &+ \mathbf{Z}_{dent}\mathbf{W}_{dent} \boldsymbol{\alpha}_{dent} + \mathbf{Z}_{dent}\mathbf{U}_{dent_{Ped}} \boldsymbol{\epsilon}_{dent_{Ped}}
+ \mathbf{Z}_{dent}\mathbf{U}_{dent_{SNP}} \boldsymbol{\epsilon}_{dent_{SNP}} \\
  &+ \mathbf{Z}_{flint}\mathbf{W}_{flint} \boldsymbol{\alpha}_{flint} + \mathbf{Z}_{flint}\mathbf{U}_{flint_{Ped}} \boldsymbol{\epsilon}_{flint_{Ped}}
+ \mathbf{Z}_{flint}\mathbf{U}_{flint_{SNP}} \boldsymbol{\epsilon}_{flint_{SNP}} \\
  &+ \mathbf{e},
\end{align}
$$

@@

$$
\begin{equation}
\tilde{\mathbf{h}} = \mathbf{Z}_{dent}\tilde{\mathbf{g}}_{dent} + \mathbf{Z}_{flint}\tilde{\mathbf{g}}_{flint} 
\end{equation}
$$

>>>>

\begin{equation}
	(\#eq:animal-model)
	\mathbf{y} = \mathbf{Xb} + \mathbf{Za} + \mathbf{e},
\end{equation}

@@

\begin{equation}
  (\#eq:gebv)
	\mathbf{\tilde{g}} =  (\mathbf{M} - 2\mathbf{p})\mathbf{\hat{u}},
\end{equation}

@@

## Single Step Transciptomic prediction of Breeding Values
The breeding values in a mRNA prediction model are in general given by:
\begin{equation}
	(\#eq:mrnaebv)
	\mathbf{\tilde{g}} = \mathbf{W}\boldsymbol{\hat{\alpha}},
\end{equation}

@@

Consider now the situation in which only a subset of the population has 
transcriptomic information but all have SNP genotypes.
We can take a similar route as in @fernando_class_2014 and impute missing
mRNA data.
The difference here is that &mdash; instead of predicting SNP covariates using 
pedigree relationship &mdash; we impute mRNA covariates using SNP or genomic 
relationships.
Let the subscript $1$ denote individuals for which SNP genotypes but no mRNA 
data are available.
Individuals with subscript $2$ have both, SNP and mRNA information.
The covariates in $\mathbf{W}$ are centered.
The vector $\mathbf{g}_1$ can be written as the sum of the conditional 
expectation given $\mathbf{g}_2$ and a residual error term:
\begin{align}
	(\#eq:mrna1)
	\mathbf{g_1} &= E(\mathbf{g}_1|\mathbf{g}_2) + \boldsymbol{\epsilon} \\
	&= \mathbf{G_{12}}\mathbf{G_{22}}^{-1}\mathbf{W_2}\boldsymbol{\hat{\alpha}} + (\mathbf{g_1} - \mathbf{G_{12}}\mathbf{G_{22}}^{-1}\mathbf{W_2}\boldsymbol{\hat{\alpha}}) \\
	&= \mathbf{\hat{g}}_1 + \boldsymbol{\epsilon}
\end{align}

@@

\begin{equation}
  (\#eq:mrna-model)
\mathbf{y} = \mathbf{Xb} + \mathbf{W} \boldsymbol{\alpha} + \mathbf{U}_{Ped} \boldsymbol{\epsilon}_{Ped} 
+ \mathbf{U}_{SNP} \boldsymbol{\epsilon}_{SNP}+ \mathbf{e},
\end{equation}

@@

\begin{equation}
 (\#eq:mrna-submatrices)
\mathbf{X} = 
\begin{bmatrix}
  X_0 \\
  X_1 \\
  X_2 
 \end{bmatrix},
 \mathbf{W} = 
\begin{bmatrix}
  Z_0\hat{\mathbf{W}_0} \\
  Z_1\hat{\mathbf{W}_1} \\
  \mathbf{W}_2 
 \end{bmatrix},
 \mathbf{U}_{Ped} = 
\begin{bmatrix}
  Z_0 \\
  0 \\
  0 
 \end{bmatrix},
 \mathbf{U}_{SNP} = 
\begin{bmatrix}
  0 \\
  Z_1 \\
  0 
 \end{bmatrix},
\end{equation}

@@

\begin{align}
 (\#eq:entangled-augmented-mrna-model)
\begin{bmatrix}
  y_0 \\
  y_1 \\
  y_2 
 \end{bmatrix}
& =
 \begin{bmatrix}
  X_0 \\
  X_1 \\
  X_2 
 \end{bmatrix}
 \boldsymbol{\beta} + 
 \begin{bmatrix}
  Z_0 & 0 & 0 \\
  0 & Z_1 & 0 \\
  0 & 0 & Z_2 
 \end{bmatrix}
\begin{bmatrix}
  g_0 \\
  g_1 \\
  g_2 
 \end{bmatrix}
  + \mathbf{e} \\
    & = 
 \begin{bmatrix}
  X_0 \\
  X_1 \\
  X_2 
 \end{bmatrix}
 \boldsymbol{\beta} + 
 \begin{bmatrix}
  Z_0 & 0 & 0 \\
  0 & Z_1 & 0 \\
  0 & 0 & Z_2 
 \end{bmatrix}
\begin{bmatrix}
  \mathbf{A}_{02}\mathbf{A}_{22}^{-1}\mathbf{W}_2\boldsymbol{\alpha} + \boldsymbol{\epsilon}_{Ped} \\
  \mathbf{G}_{12}\mathbf{G}_{22}^{-1}\mathbf{W}_2\boldsymbol{\alpha} + \boldsymbol{\epsilon}_{SNP}  \\
  \mathbf{W}_2\boldsymbol{\alpha} \\
 \end{bmatrix}
  + \mathbf{e} \\
    & = 
 \begin{bmatrix}
  X_0 \\
  X_1 \\
  X_2 
 \end{bmatrix}
 \boldsymbol{\beta} + 
 \begin{bmatrix}
  Z_0 & 0 & 0 \\
  0 & Z_1 & 0 \\
  0 & 0 & Z_2 
 \end{bmatrix}
\begin{bmatrix}
  \hat{\mathbf{W}_0}\boldsymbol{\alpha} + \boldsymbol{\epsilon}_{Ped} \\
  \hat{\mathbf{W}_1}\boldsymbol{\alpha} + \boldsymbol{\epsilon}_{SNP} \\
  \mathbf{W}_2\boldsymbol{\alpha} \\
 \end{bmatrix}
  + \mathbf{e} \\
\end{align}

@@

\begin{equation}
 (\#eq:breeding-values)
\tilde{\mathbf{g}} = 
 \begin{bmatrix}
  \hat{\mathbf{W}_0} \\
  \hat{\mathbf{W}_1} \\
  \mathbf{W}_2 
 \end{bmatrix}
 \hat{\boldsymbol{\alpha}}
 + 
 \begin{bmatrix}
  \mathbf{Z}_0 \\
  0 \\
  0
 \end{bmatrix}
 \hat{\boldsymbol{\epsilon}}_{Ped}
 + 
 \begin{bmatrix}
  0 \\
  \mathbf{Z}_1 \\
  0
 \end{bmatrix}
 \hat{\boldsymbol{\epsilon}}_{SNP} 
\end{equation}

@@

\begin{align}
 (\#eq:final-model)
\mathbf{y} &= \mathbf{Xb} \\ 
  &+ \mathbf{Z}_{dent}\mathbf{W}_{dent} \boldsymbol{\alpha}_{dent} + \mathbf{Z}_{dent}\mathbf{U}_{dent_{Ped}} \boldsymbol{\epsilon}_{dent_{Ped}}
+ \mathbf{Z}_{dent}\mathbf{U}_{dent_{SNP}} \boldsymbol{\epsilon}_{dent_{SNP}} \\
  &+ \mathbf{Z}_{flint}\mathbf{W}_{flint} \boldsymbol{\alpha}_{flint} + \mathbf{Z}_{flint}\mathbf{U}_{flint_{Ped}} \boldsymbol{\epsilon}_{flint_{Ped}}
+ \mathbf{Z}_{flint}\mathbf{U}_{flint_{SNP}} \boldsymbol{\epsilon}_{flint_{SNP}} \\
  &+ \mathbf{e},
\end{align}

@@

\begin{equation}
 (\#eq:predicted-performance)
\tilde{\mathbf{h}} = \mathbf{Z}_{dent}\tilde{\mathbf{g}}_{dent} + \mathbf{Z}_{flint}\tilde{\mathbf{g}}_{flint} 
\end{equation}

### 2016-12-07 reports/single_step_kernels.Rmd
Update equation numbering after updates to bookdown package

Goal: Inspection of input matrices for single-step hybrid prediction using 
BGLR.

>>>>

Goal: Inspection of input matrices for single-step hybrid prediction using 
BGLR.
With the original configuration of the predictors, very low predictive abilities
were obtained for the amalgamation of pedigree, genomic and transcriptomic 
information, regardless of the phenotypic trait.
One of the possible causes were inconsistencies in the functions `impute_eta()`
and `impute2` from the `sspredr` package regarding the factorization of 
matrices.
After tackling those bugs, we wanted to revisit the distribution of the 
coefficients from the matrix $\mathbf{W}$ in \@ref(eq:mrna-model).
